---
title: "Global Measles Trends"
author: "William Cuthbert, Sebastien Montgrain, Adrian Antonelli"
format:
  dashboard:
    theme: journal
    logo: "https://upload.wikimedia.org/wikipedia/commons/thumb/8/84/WHO_Logo.png/320px-WHO_Logo.png"
    nav-buttons:
      - icon: linkedin
        href: https://www.linkedin.com
        aria-label: LinkedIn
      - icon: instagram
        href: https://www.instagram.com
        aria-label: Instagram
      - icon: twitter-x
        href: https://www.twitter.com
        aria-label: Twitter/X
      - icon: envelope-fill
        href: mailto:your@email.com
        aria-label: Email
---

```{r}
#| message: false
#| include: false
library(tidyverse)
library(ggplot2)
library(forcats)
library(scales)
library(ggridges)
library(gt)
library(purrr)

cases_month <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')
cases_year  <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
complete_data <- read.csv("complete_data.csv")

country_year <- complete_data |>
  distinct(region, country, year, .keep_all = TRUE) |>
  filter(!is.na(COVERAGE_MCV1),
         !is.na(yearly_measles_incidence_rate_per_1000000_total_population))
```

# Vaccination & Incidence {scrollable="true"}

## Row {height="25%"}

::: {.card}
**Research Question 1:** How does vaccination percentage affect measles incidence rates by region?

Using immunization coverage data merged with monthly/annual case data (2012–2024), we examine whether higher MCV1 and MCV2 vaccination rates have an affect on measles incidence across the six WHO world regions.
:::

## Row {height="75%"}

::: {.panel-tabset}

### By Region

```{r}
Regional_Vaccination <- function(r) {
  country_year |>
    filter(region == r) |>
    summarise(
      region = r,
      n_countries = n_distinct(country),
      year_range = paste(min(year), 
                         max(year), sep = "–"),
      avg_mcv1_coverage = round(mean(COVERAGE_MCV1, na.rm = TRUE), 1),
      avg_mcv2_coverage = round(mean(COVERAGE_MCV2, na.rm = TRUE), 1),
      avg_incidence_per_1M = round(
        mean(yearly_measles_incidence_rate_per_1000000_total_population, na.rm = TRUE), 1))}

country_year |>
  distinct(region) |>
  pull(region) |>
  map_dfr(Regional_Vaccination) |>
  arrange(desc(avg_incidence_per_1M)) |>
  rename(
    "Region" = region,
    "Countries" = n_countries,
    "Year Range" = year_range,
    "Avg MCV1 Coverage %" = avg_mcv1_coverage,
    "Avg MCV2 Coverage %" = avg_mcv2_coverage,
    "Avg Incidence per 1M" = avg_incidence_per_1M) |>
  mutate(
    `Avg MCV1 Coverage %` = paste0(`Avg MCV1 Coverage %`, "%"),
    `Avg MCV2 Coverage %` = paste0(`Avg MCV2 Coverage %`, "%"),
    `Avg Incidence per 1M` = formatC(`Avg Incidence per 1M`, format = "f", digits = 1)) |>
  gt() |>
  tab_header(title = "Average Vaccination Coverage and Measles Incidence Rate by WHO Region (2012–2024)") |>
  tab_spanner(label = "Regional Statistics",
              columns = c(`Avg MCV1 Coverage %`, `Avg MCV2 Coverage %`, `Avg Incidence per 1M`)) |>
  tab_options(
    column_labels.font.weight = "bold",
    heading.background.color = "#2C7BB6",
    column_labels.background.color = "#D9EAF7",
    row.striping.background_color  = "#F0F7FF",
    row.striping.include_table_body = TRUE) |>
  tab_style(style = cell_text(color = "white", weight = "bold"), locations = cells_title()) |>
  tab_style(style = cell_fill(color = "#D9EAF7"),
            locations = cells_column_spanners()) |>
  tab_source_note(source_note = "Data source: World Health Organization (WHO) Measles & Rubella Surveillance, 2012–2024")
```

### General European Country analysis

```{r}
European_Data <- country_year |> filter(region == "EUR")

Country_stats <- function(c) {
  European_Data |>
    filter(country == c) |>
    summarise(country = c,
      year_range = paste(min(year), max(year), sep = "–"),
      avg_mcv1_coverage = round(mean(COVERAGE_MCV1, na.rm = TRUE), 1),
      avg_incidence_per_1M = round(
        mean(yearly_measles_incidence_rate_per_1000000_total_population, na.rm = TRUE), 1))}

European_Data |>
  distinct(country) |>
  pull(country) |>
  map_dfr(Country_stats) |>
  arrange(desc(avg_incidence_per_1M)) |>
  rename(
    "Country" = country,
    "Year Range" = year_range,
    "Avg MCV1 Coverage %"  = avg_mcv1_coverage,
    "Avg Incidence per 1M" = avg_incidence_per_1M) |>
  mutate(
    `Avg MCV1 Coverage %` = paste0(`Avg MCV1 Coverage %`, "%"),
    `Avg Incidence per 1M` = formatC(`Avg Incidence per 1M`, format = "f", digits = 1)) |>
  head(10) |>
  gt() |>
  tab_header( title = "Measles Incidence and MCV1 Coverage by Country",
    subtitle = "Europe's Top 10 Highest Incidence Countries (World's Most Vaccinated Region)") |>
  tab_spanner(label = "Country Statistics",
              columns = c(`Avg MCV1 Coverage %`, `Avg Incidence per 1M`)) |>
  tab_options(
    column_labels.font.weight = "bold",
    heading.background.color = "#1A7A4A",
    column_labels.background.color = "#D4EDDA",
    row.striping.background_color = "#F0FAF4",
    row.striping.include_table_body = TRUE) |>
  tab_style(style = cell_text(color = "white", weight = "bold"), locations = cells_title()) |>
  tab_style(style = cell_fill(color = "#D4EDDA"),
            locations = cells_column_spanners()) |>
  tab_source_note(source_note = "Data source: World Health Organization (WHO) Measles & Rubella Surveillance, 2012–2024")
```

### Top 5 European Countries over time

```{r}
#| fig-alt: "Faceted line chart of annual measles incidence per 1M for Europe's 5 highest incidence countries, 2012–2024"
#| fig-cap: "Figure 1"

top5_eur <- country_year |>
  filter(region == "EUR") |>
  group_by(country) |>
  summarise(avg_incidence = mean(yearly_measles_incidence_rate_per_1000000_total_population, na.rm = TRUE)) |>
  slice_max(avg_incidence, n = 5)

top5_data <- country_year |> filter(country %in% top5_eur$country)

ggplot(top5_data, aes(x = year,
                      y = yearly_measles_incidence_rate_per_1000000_total_population,
                  color = country, fill = country)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3, shape = 21, color = "white", stroke = 1.5) +
  facet_wrap(~ country, ncol = 2) +
  scale_x_continuous(breaks = seq(2012, 2024, by = 2)) +
  scale_y_continuous(breaks = seq(0, 3000, by = 500), labels = scales::comma) +
  scale_color_manual(values = c(
    "Georgia" = "#E63944", 
    "Kazakhstan" = "#F4A265",
    "Kyrgyzstan" = "#2A9E6D", 
    "Romania" = "#457B9D", 
    "Ukraine" = "purple")) +
  scale_fill_manual(values  = c(
    "Georgia" = "#E63944",
    "Kazakhstan" = "#F4A265",
    "Kyrgyzstan" = "#2A9E6D",
    "Romania" = "#457B9D",
    "Ukraine" = "purple")) +
  labs(title = "Measles Outbreaks Over Time in Europe's 5 Highest Incidence Countries",
       subtitle = "Annual measles incidence per 1,000,000 population (2012–2024)",
       x = "Year", 
       y = "Incidence per 1,000,000 Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14),
        strip.text = element_text(face = "bold", size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8), legend.position = "none")
```

### MCV1 Coverage % vs. Incidence per 1M 

```{r}
#| fig-alt: "Density-ridge plots showing measles incidence distributions by region, colored by average MCV1 coverage"
#| fig-cap: "Figure 2"
#| message: false

disease_ridges <- complete_data |>
  filter(!is.na(COVERAGE_MCV1), year <= 2024,
         yearly_measles_incidence_rate_per_1000000_total_population > 0) |>
  select(region, year, country, COVERAGE_MCV1,
         yearly_measles_incidence_rate_per_1000000_total_population) |>
  mutate(avg_mcv1 = mean(COVERAGE_MCV1), .by = region)

ggplot(disease_ridges,
       aes(x = yearly_measles_incidence_rate_per_1000000_total_population,
           y = region, fill = avg_mcv1)) +
  geom_density_ridges(alpha = 0.8, scale = 1.5, quantile_lines = TRUE, quantiles = 2) +
  geom_text(data = disease_ridges |> distinct(region, avg_mcv1),
            aes(x = 1075, y = region, label = paste0("MCV1: ", round(avg_mcv1), "%")),
            vjust = -1.5, size = 3) +
  scale_x_log10(labels = comma, breaks = c(0.1, 1, 10, 100, 1000), minor_breaks = NULL) +
  scale_fill_gradient(low = "dodgerblue", high = "darkred", name = "Avg MCV1 (%)") +
  labs(title = "Do regions with higher avg MCV1 vaccination percentages have lower measles cases?",
       subtitle = "Vertical line represents the median of the region.",
       x = "Measles incidence rate per 1 million population", 
       y = NULL,
       caption  = "AFR = Africa | AMR = Americas | EMR = E. Mediterranean | EUR = Europe | SEAR = SE Asia | WPR = W. Pacific") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(), legend.position = "none")
```

:::

# Seasonality {scrollable="true"}

## Row {height="25%"}

::: {.card}
**Research Question 2:** Does time of year impact measles cases globally?

Using monthly case counts across all WHO regions, we are interested in finding whether measles incidence is affected by seasonal patterns.
:::

## Row {height="75%"}

::: {.panel-tabset}

### Monthly Cases by Region

```{r}
#| fig-alt: "Table showing average measles cases by month for each world region"

case_ranking_by_months <- cases_month |>
  group_by(month, region) |>
  filter(measles_total >= 0) |>
  summarise(avg_measles = mean(measles_total, na.rm = TRUE), .groups = "drop") |>
  mutate(month = factor(month.abb[month], levels = month.abb),
         region = recode(region,
                         "AFR"  = "Africa",
                         "AMR"  = "Americas",
                         "EMR"  = "Eastern Mediterranean",
                         "EUR"  = "Europe",
                         "SEAR" = "South-East Asia",
                         "WPR"  = "Western Pacific"))

case_ranking_by_months |>
  pivot_wider(names_from = "month", values_from = "avg_measles") |>
  gt() |>
  fmt_number(columns = where(is.numeric), decimals = 2) |>
  cols_label(region = "Region") |>
  tab_header(title = "Average Monthly Measles Cases by Region") |>
  tab_style(style = cell_text(weight = "bold"), locations = cells_title(groups = "title")) |>
  tab_source_note(source_note = "Data source: World Health Organization (WHO)")
```

### Monthly Trend Lines

```{r}
#| fig-alt: "Line plot showing average measles cases by month per region with trend lines"
#| fig-cap: "Figure 3"
#| message: false

ggplot(case_ranking_by_months,
       aes(x = month,
                  y = avg_measles, 
                  group = region, 
                  color = region)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5, linetype = "dashed", col = "black") +
  facet_wrap(~region, scales = "free_y") +
  labs(x = "Month", title = "Measles Cases by Month Across the World", y = "",
       subtitle = "Number of Total Measles Cases (Clinical, EPI linked, and Lab Confirmed)") +
  theme_linedraw() +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(size = 6), legend.position = "none")
```

### Seasonal Summary

```{r}
#| fig-alt: "Table displaying average monthly measles cases by season"

largest_countries <- c("India", "China", "United States of America","Indonesia", "Pakistan", "Nigeria")

season_lookup <- tribble(
  ~country, ~month, ~season,
  "India", "Apr", "Summer", "India", "May", "Summer", "India", "Jun", "Summer",
  "India", "Dec", "Winter", "India", "Jan", "Winter", "India", "Feb", "Winter",
  "China", "Jun", "Summer", "China", "Jul", "Summer", "China", "Aug", "Summer",
  "China", "Dec", "Winter", "China", "Jan", "Winter", "China", "Feb", "Winter",
  "United States of America", "Jun", "Summer", "United States of America", "Jul", "Summer",
  "United States of America", "Aug", "Summer", "United States of America", "Dec", "Winter",
  "United States of America", "Jan", "Winter", "United States of America", "Feb", "Winter",
  "Indonesia", "Sep", "Summer", "Indonesia", "Oct", "Summer", "Indonesia", "Nov", "Summer",
  "Indonesia", "Jan", "Winter", "Indonesia", "Feb", "Winter", "Indonesia", "Mar", "Winter",
  "Pakistan", "May", "Summer", "Pakistan", "Jun", "Summer", "Pakistan", "Jul", "Summer",
  "Pakistan", "Dec", "Winter", "Pakistan", "Jan", "Winter", "Pakistan", "Feb", "Winter",
  "Nigeria", "Feb", "Summer", "Nigeria", "Mar", "Summer", "Nigeria", "Apr", "Summer",
  "Nigeria", "Jul", "Winter", "Nigeria", "Aug", "Winter", "Nigeria", "Sep", "Winter")

get_season <- function(country_name, month_name) {
  month_name <- as.character(month_name)
  out <- season_lookup |> filter(country == country_name, month == month_name) |> pull(season)
  if (length(out) == 0) return("Other") else return(out[1])
}

country_month_summary <- cases_month |>
  filter(country %in% largest_countries) |>
  group_by(month, country) |>
  summarise(avg_measles = mean(measles_total, na.rm = TRUE), .groups = "drop") |>
  mutate(month = factor(month.abb[month], levels = month.abb),
         season = map2_chr(country, as.character(month), get_season))

country_month_summary |>
  group_by(season) |>
  summarize(avg_measles = mean(avg_measles)) |>
  gt() |>
  fmt_number(columns = avg_measles, decimals = 2) |>
  cols_label(season = "Season", avg_measles = "Average Measles Cases") |>
  tab_header(title = "Average Monthly Measles Cases per Country by Season") |>
  tab_style(style = cell_text(weight = "bold"), locations = cells_title(groups = "title")) |>
  tab_source_note(source_note = "Data from most populated country from each region (six countries total)")
```

### Seasonal Distribution

```{r}
#| fig-alt: "Violin plot showing distribution of measles cases by season for six representative countries"
#| fig-cap: "Figure 4"

ggplot(country_month_summary, aes(x = season, y = avg_measles, fill = season)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  labs(title    = "Average Measles Cases per Country by Season",
       subtitle = "Summer and winter determined by average temperature, data from six representative countries",
       x = "Season", y = "") +
  scale_fill_manual(values = c("Summer" = "darkred", "Winter" = "dodgerblue", "Other" = "#1A7A4A")) +
  theme_bw() +
  theme(legend.position = "none")
```

### Statistical Tests

```{r}
#| fig-alt: "Table displaying F-test results for month and season predictors"

tibble(
  Model = c("Month", "Season"),
  Predictor = c("Month", "Season"),
  DF_numerator = c(1, 2),
  DF_denominator = c(22630, 69),
  F_statistic = c(68.34, 0.658),
  p_value = c("< 2e-16", "0.521"),
  Conclusion = c("Significant", "Not significant")) |>
  gt() |>
  fmt_number(columns = F_statistic, decimals = 2) |>
  cols_label(DF_numerator = "DF Numerator", DF_denominator = "DF Denominator",
             F_statistic = "F Statistic", p_value = "p-value") |>
  tab_header(title = "Testing for Significance: Month and Season as Predictors of Measles Cases") |>
  tab_style(style = cell_text(weight = "bold"), locations = cells_title(groups = "title"))
```

:::

# About

## Row

::: {.card}
### About This Dashboard

We are focused on researching measles case trends between the years of 2012-2025 and we are interested in the following topics:

1. **Vaccination & Incidence** — How does vaccination percentages affect measles incidence rates by region?

2. **Seasonality** — Does time of year impact measles cases globally?


**Authors:** William Cuthbert · Sebastien Montgrain · Adrian Antonelli
:::
